---
title: "Extended Analysis of Thermal Tolerance Landscapes"
author: "Andrew Villeneuve"
date: "2024-04-12"
output: html_document
---

# Readme

This script was written to construct the extended heatmap in the supplementary material of Villeneuve and White 2024 MHW simulation paper. This markdown is meant to be a stand-alone document from the main analysis of the paper (mhwsim_analysis_threesp.Rmd) which can be knitted. Note, however, that a bulk of the analysis is silenced here due to the extreme computational requirements of analyzing more than double the MHW profiles from the original, main text analysis. We ran this code on the University of New Hampshire's supercomputing cluster PREMISE, which took around six hours to complete. We highly suggest running the code found in the main RMD document unless the intention is to completely replicate our results, in which case the first code chunk should be unsilenced. The R code for running this analysis on a computing cluster is available upon request to Andrew Villeneuve. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(lubridate)
require(tibble)
require(ggplot2)
require(purrr)
require(ggridges)
require(parallel)
require(foreach)
require(data.table)
require(dtplyr)
require(zoo)
require(cowplot)
library(RColorBrewer)
require(doParallel)
require(rlang)
require(reshape2)
require(heatwaveR)
require(ggpubr)
require(SpatialTools)


options(scipen=999)

source("Thermal_landscape_functions.R")
source("hour_fxns_heatwaveR.R")

    #a function to turn NAs in a data table into zeros (useful for binding datatables of different lengths)
f_dowle3 = function(DT) {
  for (j in seq_len(ncol(DT)))
    set(DT,which(is.na(DT[[j]])),j,0)
}
```

```{r,eval=F}
##################################################
magnitude_min<-0.01 #0.01 degree minimum magnitude. Analysis does not accept 0 value
magnitude_max<-20.01 #8 degree duration max. This is in agreement with observed values above climatology across multiple major MHW events
duration_min<-60*1 #one hour MHW min  Here, we differ from other definitions (Hobday et al. 2016) that set minimum values of 5 days for a MHW event. For intertidal organisms, a few hours of extreme heat during low tide during an atmospheric heatwave may cause mortality and should be considered from an ecophysiological perspective. 
duration_max<-24*30*60+60*1 # 30 day MHW max - MHWs can commonly last multiple days, weeks, or months. However, for the sake of this initial analysis, we reduce that to three day for speed of calculation.Note that this is technically 30 days plus one hour, to keep intervals between durations the same. 

magnitude_inc<-0.5 # magnitudes given in 0.5 degree increments 
duration_inc<-24*60 #durations given in 24 hour increments (MUST be in minutes)
slope_inc<-NA

MHW_tri<-function(magnitude_min,magnitude_max,duration_min,duration_max,magnitude_inc,duration_inc,slope_inc){
  
  #sequence of magnitudes
  mag_seq<-seq(magnitude_min,magnitude_max,by = magnitude_inc)
  
  #sequence of duration (in minutes)
  dur_seq<-seq(duration_min,duration_max,by = duration_inc)
  
  if (is.na(slope_inc)){
    
    #create an expanded dataframe of all possible parameter combinations. 
    
    expanded<-data.frame(expand.grid('mag_seq'=mag_seq,'dur_seq'=dur_seq))
  } else{
    slope_seq<-seq(0.0001,6/360,by=slope_inc)
    expanded<-data.frame(expand.grid('mag_seq'=mag_seq,'dur_seq'=dur_seq,'slope_seq'=slope_seq)) # if we want to control slope
  }
  
  
  
  
  
  #create empty matrix to store results
  mag_dur_tri <- vector("list", length(mag_seq))
  
  #single loop to create triangles for each row in the expanded dataframe, encapsulating all possible parameters  
  for (i in 1:nrow(expanded)){
    magnitude<-as.numeric(expanded[i,1])
    hw_length<-expanded[i,2] #duration
    #rising_slope<-expanded[i,3] if we vary slope
    rising_slope<-magnitude/(hw_length/2)
    #create df of dates and numerical sequence
    
    rising_length<-magnitude/rising_slope
    falling_length<-hw_length-rising_length
    if (is.na(rising_length) || is.na(hw_length)) {
      print("Debug: Missing value in rising_length or hw_length")
      #add ifelse to remove unlikely initial slope values
    } else if(rising_length>=hw_length){#if rising portion is longer than entire mhw duration
      z<-data.frame("timestep"=NA,"value"=NA,"magnitude"=magnitude,"duration"=hw_length,"rising_slope"=rising_slope,"falling_slope"=NA,"area"=NA,"mean_i"=NA,"hdd"=NA)
    } else {
      falling_slope<-(-rising_slope)
      #dataframes of rising falling points
      rising_seq<-data.frame("timestep"=seq(1,floor(max(rising_length)),by=1))
      falling_seq<-data.frame("timestep"=seq(from=floor(max(rising_length))+1,to=hw_length,by = 1))
      
      rising_seq$value<-rising_seq$timestep*rising_slope
      falling_seq$value<-(falling_seq$timestep*falling_slope)+(max(rising_length)*magnitude/nrow(falling_seq))+magnitude
      
      #apply values
      
      z<-rbind(rising_seq,falling_seq)
      z$reps<-i
      z$magnitude<-magnitude
      z$duration<-hw_length
      z$rising_slope<-rising_slope
      z$falling_slope<-falling_slope
      z$area<- magnitude*hw_length/2
      z$mean_i<-mean(z$value)
      #heating degree days
      #first, create a column of day 
      if(hw_length<1440){
        z$unique_day<-hw_length
      }else{
        z$unique_day<-(z$timestep-1)%/%1440+1
      }
      
      hdd_tib<-z%>%group_by(unique_day)%>%summarize(mean_t=mean(value))%>%summarize(output=sum(mean_t))
      z$hdd<-hdd_tib$output
    }
    
    
    mag_dur_tri[[i]] <- z
  }
  
  boundoutput_tri<-bind_rows(mag_dur_tri,.id="reps") #from matrix to data.frame
  boundoutput_tri$reps<-as.numeric(boundoutput_tri$reps)
  saveRDS(boundoutput_tri,"data/boundoutput_tri_ext.RDs") #saves file to onedrive due to space restrictions in github. Can be silenced. 
  

}

system.time(MHW_tri(magnitude_min,magnitude_max,duration_min,duration_max,magnitude_inc,duration_inc,slope_inc))

tri_output<-readRDS("data/boundoutput_tri_ext.RDs")
tri_data<-tri_output
tri_data$duration<-tri_data$duration/60 #convert to hours for plotting

tri_data$label<-paste(tri_data$magnitude, "°C", sep = "")
tri_data<-as.data.frame(tri_data)
waves <- function(date_start,date_end, alpha = 0, beta = 1, freq = 24, phi = 0){
  
  start <- as.POSIXct(date_start)
  end<-as.POSIXct(date_end)
  x <- seq(start, end, by = 60 * 1)
  
  
  # timestep per hour
  time_step <- 60 / unique(diff(x))
  # set phi as difference in hours from start of time_in
  phi <- min(x) + phi * 3600
  phi<- as.numeric(difftime(phi, min(x)))
  phi <- phi / time_step
  # get input values to cos func
  in_vals <- seq(0, length(x), length = length(x))
  in_vals <- in_vals / time_step
  in_vals <- 2 * pi * in_vals * 1 / freq
  # wave
  y <- alpha + beta * sin((in_vals + phi)-(pi*.75))
  df<-data.frame("vals"=y,"datetime"=x)
  return(df)
}

year_out<-waves('2022-01-01','2022-12-31',alpha=12,beta=14,freq=365*24)

#list of range parameters
alpha_list<-c(12.75)
merged_temp_tri_alphas<-vector("list",length=length(alpha_list))

system.time(
  for (xx in 1:length(alpha_list)){
    year_thresh<-waves('2022-01-01','2022-12-31',alpha=alpha_list[xx],beta=14,freq=365*24)
    
    
    clim<-data.frame("hoy"=lubridate::hour(year_out$datetime) + (lubridate::yday(year_out$datetime) - 1) * 24,
                     "doy"=lubridate::yday(year_out$datetime),
                     "t"=year_out$datetime,
                     "temp"=year_out$vals,#climatology
                     "seas"=year_out$vals,
                     "thresh"=year_thresh$vals,
                     "thresh_2x"=(year_thresh$vals-year_out$vals)+year_thresh$vals,
                     "thresh_3x"=(((year_thresh$vals-year_out$vals)+year_thresh$vals)-year_out$vals)+year_thresh$vals,
                     "thresh_4x"=(((((year_thresh$vals-year_out$vals)+year_thresh$vals)-year_out$vals)+year_thresh$vals)-year_out$vals)+year_thresh$vals)
    
    
    clim_2022<-clim%>%dplyr::filter(t>=as.POSIXct('2022-07-22')&t<as.POSIXct('2022-09-12'))
    out_res<-clim_2022%>%dplyr::rename("datetime"=t,"vals"=temp)
    merged_temp_tri<-vector("list",length(unique(tri_data$reps)))
    maxtemptime<-out_res%>%dplyr::filter(datetime>as.POSIXct('2022-01-01 00:00:00'))%>%
      dplyr::filter(vals==max(vals))%>%dplyr::select(datetime)
    maxtemptime<-as.POSIXct(maxtemptime[,1])
    reps_list<-unique(tri_data$reps)
    length(reps_list)
    
    #user 2238 system 160 elapsed 3439  - long, but better than an unoptimized left_join at 14000 seconds
    #time 2, takes 10684s with an additional 22 simulations
    syst<-system.time(
      for (i in 1:length(reps_list)){
        tryCatch({
        #setup
        rep_i<-tri_data%>%dplyr::filter(reps==i) # filter for each MHW
        start_date_rep<-maxtemptime-floor(max(rep_i$duration/2))*60 #find halfway point and subtract from maximum annual timestamp to give start time
        end_date_rep<-start_date_rep+ceiling(max(rep_i$duration))*60
        rep_i$datetime<-seq(from=as.POSIXct(start_date_rep),length.out=ceiling(max(rep_i$timestep)),by="min")
        
        #much faster method of left_join and na.replace
        joined_df<-as.data.table(rep_i)[as.data.table(out_res),on='datetime']
        system.time(f_dowle3(joined_df)) # another solution
        
        joined_df<-as.data.frame(joined_df)
        joined_df$value<-ifelse(is.na(joined_df$value),0,joined_df$value)
        joined_df$combined_temperature <- joined_df$vals + joined_df$value
        joined_df_short<-joined_df%>%dplyr::filter(datetime>=as.POSIXct('2022-07-22 00:00:00')&datetime<as.POSIXct('2022-09-12 00:00:00'))
        joined_df_short<-joined_df_short%>%rename("time"=datetime,"temp"=combined_temperature)
        
        #now, we categorize each MHW according to the Hobday definition. Note that we have altered code to 1) allow for timeseries in hours (and thus heat wave events classified in hours) and 2) we have removed minimum MHW length (previosuly five days)
        repi_selection<-rep_i%>%dplyr::select(c(value,datetime))%>%dplyr::rename("t"=datetime)%>%dplyr::filter(t>='2022-07-22 00:00:00'&t<'2022-09-12 00:00:00')
        
        #0.11
        result_df <- clim_2022 %>%
          left_join(repi_selection, by = "t") %>%
          mutate(sum_temp_value = coalesce(temp + value, temp))%>%dplyr::select(-c(temp,value))%>%dplyr::rename("temp"=sum_temp_value)
        
        
        events<-detect_event_hour(result_df,date_start='2022-07-22',date_end='2022-09-12')
        event_categories<-heatwaveR::category(events,S=F,name="simulation")
        
        #because the thresholds are not smoothed function, we occasionally get 
        if (nrow(event_categories)>1){
          event_categories<-event_categories%>%dplyr::slice(which.max(duration))
        }
        
        joined_df_short$reps<-reps_list[i]
        joined_df_short$duration_hob<-event_categories$duration
        joined_df_short$peak_date<-event_categories$peak_date
        joined_df_short$category<-event_categories$category
        joined_df_short$imax_hob<-event_categories$i_max
        joined_df_short$p_moderate<-event_categories$p_moderate
        joined_df_short$p_strong<-event_categories$p_strong
        joined_df_short$p_severe<-event_categories$p_severe
        joined_df_short$p_extreme<-event_categories$p_extreme
        

        
        merged_temp_tri[[i]]<-joined_df_short
        rm(joined_df_short)
        })
              }
      
    )
    saveRDS(merged_temp_tri,paste0("data/merged_temp_tri_",xx,".RDs"))
    merged_temp_tri<-purrr::compact(merged_temp_tri)
    
    merged_temp_tri_alphas[[xx]]<-merged_temp_tri
  }
)

saveRDS(merged_temp_tri_alphas,"data/merged_temp_tri_alphas_ext.RDs")
merged_temp_tri_alphas<-readRDS("data/merged_temp_tri_alphas_ext.RDs")

try<-merged_temp_tri_alphas[[1]][[1]]

#################################################
set.seed(123)

#Let's create three species: Two of which have intersecting TDT at 30C and 15 days. The third will intersect the high CT species at 30C at 1 day. We will set CTmax

#first, find potential values of CTmax to work with that intsersect at 15 days at 30C
temp.seq<-seq(35,50,by=1)
z_list<-vector("list",length=length(temp.seq))

for (i in 1:length(temp.seq)){
  slope<-(log10(15*60*24-1)/(30-temp.seq[i]))
  z_list[[i]]<-(-1/slope)
}

z_bound<-unlist(z_list)
sim_vals<-data.frame("ctmax"=temp.seq,"z"=z_bound)

#we select 48C and 44C to start off
highct_highz<-sim_vals[14,]
highct_highz$species<-"highct_highz"
lowct_midz<-sim_vals[10,]
lowct_midz$species<-"lowct_midz"

#thus far, we have selected tdt curves that intersect in thè middle of our parameter field (15 days, 4C above)
#we now are interestd in a species that still has a lower CTmax, but will have a lower z such that it intersects the highCTmax Highz species quickly at 12 hours.

equaltdt <- function(CTmax1, CTmax2, z1, z2, timetodeath) {
  if (is.na(CTmax2)) {
    CTmax2 <- CTmax1 - (log10(timetodeath) * (z1 - z2))
    tempatdeath <- CTmax2 - z2 * log10(timetodeath)
    return(list(CTmax2 = CTmax2, z2 = z2, temp_at_equal_TDT = tempatdeath))
  } else if (is.na(z2)) {
    z2 <- z1 - ((CTmax1 - CTmax2) / log10(timetodeath))
    tempatdeath <- CTmax2 - z2 * log10(timetodeath)
    return(list(z2 = z2, temp_at_equal_TDT = tempatdeath))
  } else if (is.na(timetodeath)) {
    timetodeath<-10^((CTmax1-CTmax2)/(z1-z2))
    tempatdeath <- CTmax2 - z2 * log10(timetodeath)
    
    return(list(timetodeath=timetodeath,tempatdeath=tempatdeath))
  }
}

#acute-chronic
equ_res1a<-equaltdt(CTmax1=highct_highz$ctmax,CTmax2=lowct_midz$ctmax,z1=highct_highz$z,z2=NA,timetodeath=12*60)

lowct_lowz<-data.frame("ctmax"=44,"z"=equ_res1a$z2,"species"="lowct_lowz")
equ_res1b<-equaltdt(CTmax1=lowct_midz$ctmax,CTmax2=highct_highz$ctmax,z1=lowct_midz$z,z2=highct_highz$z,timetodeath=NA)


#acute-mixed
equ_res2<-equaltdt(CTmax1=lowct_lowz$ctmax,CTmax2=lowct_midz$ctmax,z1=lowct_lowz$z,z2=lowct_midz$z,timetodeath=NA)
#chronic-mixed
equ_res3<-equaltdt(CTmax1=highct_highz$ctmax,CTmax2=lowct_lowz$ctmax,z1=highct_highz$z,z2=lowct_lowz$z,timetodeath=NA)


#set ctmax and z parameters
example_TDT<-bind_rows(highct_highz,lowct_midz,lowct_lowz)


temp<-rep(seq(26,46,by=2))

# Create an empty data frame to store the results
sim_pred <- vector("list",length=nrow(example_TDT))
error<-10^(-1)

# Loop over the species
for(m in 1:nrow(example_TDT)){
  # Model of TDT curve
  time <- 10^((example_TDT$ctmax[m] - temp) / example_TDT$z[m])
  
  fit <- lm(log10(time) ~ temp)
  
  
  
  sim_results <- data.frame(species = rep(example_TDT$species[m], length(time)),
                            temp = temp,
                            time = time,
                            geom = "line")
  simulated_data <- data.frame()
  
  for (j in seq_along(temp)) {
    # Generate 10 replicates for each temperature with normal distribution
    replicates <- data.frame(
      temp = rep(temp[j], times = 10),
      time = 10^(rnorm(10, mean(predict(fit, newdata = data.frame(temp = temp[j]))), sqrt(error)))
    )
    # Add to the simulated_data
    replicates$species <- rep(example_TDT$species[m], nrow(replicates))
    replicates$geom="point"
    simulated_data <- bind_rows(simulated_data, replicates)
    
  }
  
  sim_pred[[m]] <- rbind(simulated_data, sim_results)
}

sim_pred_bound<-bind_rows(sim_pred)

simulated_line<-sim_pred_bound%>%dplyr::filter(geom=="line")
simulated_point<-sim_pred_bound%>%dplyr::filter(geom=="point")

temp.df<-simulated_point%>%dplyr::select(-"geom")


#we want both examples to have almost no mortality at no MHW
unique_simsp<-unique(sim_pred_bound$species)


models <- by(simulated_point, simulated_point$species, function(sub_df) lm(log10(time) ~ temp, data = sub_df))



merged_temp_tri<-merged_temp_tri_alphas[[1]]

rm(merged_temp_tri_alphas)
#   user  system elapsed 
#  71.71   60.69 16739  


unique_simsp<-unique(temp.df$species)
final_surv_crosssim_species<-vector("list",length(unique_simsp))
r<-vector("list",length(merged_temp_tri))

cl <- makeCluster(2)
registerDoParallel(cl)

sys.tm <- system.time({
  for (yy in 1:length(unique_simsp)) {
    temp.df_sp <- temp.df %>% dplyr::filter(species == unique_simsp[yy])
    
    results <- foreach::foreach(i = 1:length(merged_temp_tri), .packages = c('tidyverse', 'ggplot2'), .combine = 'rbind') %dopar% {
      tmpi <- rezende_min(merged_temp_tri[[i]], temp.df_sp,tc=26)
      temp_plat_rep <- purrr::map_dfr(seq_len(2), ~merged_temp_tri[[i]])
      r_subset <- cbind(tmpi, temp_plat_rep)
      
      r_subset <- r_subset %>% dplyr::filter(type == "cum.mort")
      
      final_surv_crosssim_species_n <- data.frame(
        'cum.mort' = min(na.omit(r_subset$mort)),
        "magnitude" = max(r_subset$magnitude,na.rm=T),
        "duration" = max(r_subset$duration,na.rm=T),
        "rising_slope" = max(r_subset$rising_slope,na.rm=T),
        "falling_slope" = max(r_subset$falling_slope,na.rm=T),
        "area" = max(r_subset$area,na.rm=T),
        "rep" = max(r_subset$rep,na.rm=T),
        "mean_i" = max(r_subset$mean_i,na.rm=T),
        "hdd" = max(r_subset$hdd,na.rm=T),
        "duration_hob" = max(r_subset$duration_hob,na.rm=T),
        "peak_date" = max(r_subset$peak_date,na.rm=T),
        "category" = max(r_subset$category,na.rm=T),
        "imax_hob" = max(r_subset$imax_hob,na.rm=T),
        "p_moderate" = max(r_subset$p_moderate,na.rm=T),
        "p_strong" = max(r_subset$p_strong,na.rm=T),
        "p_severe" = max(r_subset$p_severe,na.rm=T),
        "p_extreme" = max(r_subset$p_extreme,na.rm=T),
        "species" = unique_simsp[yy],
        "alpha" = 12.75
      )
      
      return(final_surv_crosssim_species_n)
    }
    
    final_surv_crosssim_species[[yy]] <- results
  } # loop for each species
})

parallel::stopCluster(cl)
sys.tm

#27000 seconds, two parallel processes, supercomputer use



saveRDS(final_surv_crosssim_species,"data/final_surv_crosssim_species_try_ext.RDs")
#25772

```

```{r,echo=F,warning=F}
final_surv_crosssim_species<-readRDS("data/final_surv_crosssim_species_try_ext.RDs")

ext_bound<-bind_rows(final_surv_crosssim_species)


ext_bound$hours_dur<-ext_bound$duration/60

ext_bound$label <- ifelse(ext_bound$species == "highct_highz", "Acute Tolerator",
                                      ifelse(ext_bound$species == "lowct_midz", "Mixed Strategy",
                                             "Chronic Tolerator"))
ext_bound<-ext_bound%>%filter(magnitude<=20.1)
#add points of intersection
points_data <- data.frame(
  duration = c(30, 30, 1/60, 1/60,15*24,15*24),
  magnitude = c(36.1-26, 36.1-26, 44-26, 44-26,30-26,30-26),
  label = c("Acute Tolerator", "Chronic Tolerator", "Mixed Strategy", "Chronic Tolerator","Acute Tolerator","Mixed Strategy"),
  colour = c("#377EB8", "#E41A1C", "#4DAF4A", "#E41A1C","#377EB8","#4DAF4A"),
  shape = c(15, 15, 17, 17,16,16),
  intersection = c("Acute-Chronic Tolerator", "Acute-Chronic Tolerator", "Chronic Tolerator-Mixed", "Chronic Tolerator-Mixed","Acute Tolerator-Mixed","Acute Tolerator-Mixed")
)

point_col<-brewer.pal(name="Set1",n=3)



ggplot(data = ext_bound, aes(x = duration, y = magnitude)) +
  geom_tile(aes(fill = cum.mort)) +
  facet_wrap(~label) +
  scale_fill_viridis_c() +
  theme_classic() +
  labs(x = "Duration (hours)", y = "Magnitude (°C)", fill = "Survival (%)") +
  geom_rect(aes(xmin=0,xmax=30*24,ymin=0-.22,ymax=8+.22),fill=NA,inherit.aes=F,linewidth=1,color="black",linetype="dashed")+
  #geom_contour(aes(z = cum.mort),color="white")+ geom_text_contour(aes(z = cum.mort),color="white",fontface="bold")+
    geom_point(data = points_data, aes(x = duration, y = magnitude,color = colour, shape = intersection),size=2,inherit.aes=F)+
  #scale_color_manual(values = c("#377EB8", "#E41A1C", "#")) +
    scale_color_manual(values = c("#E41A1C", "#4DAF4A", "#377EB8")) +

  scale_shape_manual(name = "TDT Curve Intersection",
                     labels = c("Acute-Chronic Tolerator", "Chronic Tolerator-Mixed Strategy","Acute Tolerator-Mixed Strategy"),
                     values = c(15, 17,16))+
  guides(color="none")


ggsave(path = "result_figs/supp",filename="species_combinedplots_three_upper.tif", device='tiff', dpi=300,width=12,height=4,units="in")
ggsave(path = "result_figs/supp",filename="species_combinedplots_three_upper.pdf", device='pdf', dpi=300,width=12,height=4,units="in")
